Generative AI for image reconstruction in Intensity Interferometry: a first attempt

### 2025-11-11 Phase_II/draft/ParaGan.tex
Edited ParaGan.tex with TikZ output in pdf format
upto section 3

\section{Network Parameters}
\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{fig/testing_image/image_7.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_12.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_13.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_29.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_39.png}
\caption{Example results of image reconstruction using the GAN model along with the II observations simulated in this work. Each row in this figure represents the results for a hypothetical fast-rotating star. Going from left to right in each row, the first panel represents the simulated $(u,v)$ plane II signals obtained using the six baselines (right panel of of fig.\ref{fig:ft}). This image is presented, as the input, to the trained GAN to produce a predicted image. The second panel is the real image of the star, also called the ground truth. The third panel is the reconstructed image, or the predicted image, produced by the trained GAN model . The fourth and the last panel is the difference between the ground truth and the predicted image in the $(u,v)$ plane. The uniform background color in the difference panel suggests that the network is picking up white noise from the $(u,v)$ plane. }
	\label{fig:GAN}
\end{figure*}
Here, we discuss the parameters of the GAN architecture used for reconstructing images of stellar objects using II. Given the adversarial nature of GANs, where the Generator and Discriminator engage in a min-max game, careful tuning of key parameters is critical to ensure that both networks are well-balanced for effective training.

@@

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[node distance=0.9cm]
%\begin{tikzpicure}[scale = 0.5]

@@

\end{tikzpicture}
\end{figure}

%\end{document}

>>>>

\section{Network Parameters}
\begin{figure*}
	\centering
	\includegraphics[width=0.785\textwidth]{fig/testing_image/image_7.png}
	\includegraphics[width=0.785\textwidth]{fig/testing_image/image_12.png}
	\includegraphics[width=0.785\textwidth]{fig/testing_image/image_13.png}
	\includegraphics[width=0.785\textwidth]{fig/testing_image/image_29.png}
	\includegraphics[width=0.785\textwidth]{fig/testing_image/image_39.png}
\caption{Example results of image reconstruction using the GAN model along with the II observations simulated in this work. Each row in this figure represents the results for a hypothetical fast-rotating star. Going from left to right in each row, the first panel represents the simulated $(u,v)$ plane II signals obtained using the six baselines (right panel of of fig.\ref{fig:ft}). This image is presented, as the input, to the trained GAN to produce a predicted image. The second panel is the real image of the star, also called the ground truth. The third panel is the reconstructed image, or the predicted image, produced by the trained GAN model . The fourth and the last panel is the difference between the ground truth and the predicted image in the $(u,v)$ plane. The uniform background color in the difference panel suggests that the network is picking up white noise from the $(u,v)$ plane. }
	\label{fig:GAN}
\end{figure*}

Here, we discuss the parameters of the GAN architecture used for reconstructing images of stellar objects using II. Given the adversarial nature of GANs, where the Generator and Discriminator engage in a min-max game, careful tuning of key parameters is critical to ensure that both networks are well-balanced for effective training.

@@

\iffalse
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[node distance=0.9cm]
%\begin{tikzpicure}[scale = 0.5]

@@

\end{tikzpicture}
\end{figure}
\fi

\begin{figure}
\begin{center}
\includegraphics[scale=0.4]{FlowDiagram_TikZ.pdf}
\caption{A block diagram for the process of image reconstruction through C-GAN.}
\label{fig:FlowDiagram}
\end{center}
\end{figure}
%\end{document}

### 2025-11-11 Phase_II/draft/intro_reformatted.tex
upto section 3

The foundational basis of II stems from the pioneering experiments and theoretical investigations, widely referred to as the \textquotedblleft HBT effect \textquotedblright, carried out initially by Hanbury Brown and Twiss (HBT) \cite{HBT56Lab,brown1957interferometry, brown1958interferometry} and, later, by Glauber \citep{glauber1963quantum}. HBT reported \cite{HBT56Lab} in 1956, the correlation between photons measured by a pair of photon detectors in two coherent beams of light. Rai et al. \cite{Rai2025} present a modern-day variant of this experiment, carried out with pseudo-thermal light. The HBT Effect and the related theoretical investigations laid the foundation for the modern field of Quantum Optics.

@@

As it stands today, studies grappling various issues of image reconstruction are being reported \citep{Haubois2009, Norris2021AZCyg, Liu2024SuperresolutionII, Liu2025}. MI-based image reconstruction has made substantial progress in this area with efforts at generating constructed images of stars like Betelgeuse \citep{Haubois2009} and AZ Cyg \citep{Norris2021AZCyg}. On the other hand, II-based methods are in a nascent stage. The recent publications of Liu et al. \cite{Liu2024SuperresolutionII, Liu2025} have demonstrated, through outdoor experiments, imaging millimeter-scale targets at 1.36 km with a resolution ~14 times better than a single telescope’s diffraction limit. A \textquotedblleft flexible computational algorithm \textquotedblright reconstructs images from intensity correlations, overcoming atmospheric turbulence and optical imperfections.

We report here, the results of our attempt, the first of its kind, to reconstruct the II-simulated gravity-darkened images of fast rotating stars using a C-GAN neural network architecture. Image reconstruction in gravity darkened fast rotating stars has long been examined using various methods in MI \citep{vanBelle2001, DomicianodeSouza2003, DomicianodeSouza2005, mcalister2005first, Monnier2007, Pedretti2009, Zhao2009, Martinez2021}. Recently photosphere oblateness of $\gamma$-Cassiopeia \citep{Archer-arXiv-2025} has been measured at the VERITAS observatory using II. These results, and especially, that by \cite{Archer-arXiv-2025} put our work in context. Thus our simulation is a natural next step of the work of Archer et al and others. We emulate the C-GAN model of Isola et al. \cite{isola2017image} to reconstruct images of fast-rotating stars using their simulated Intensity Interferograms and simulated sky-intensity distributions as input data for training, testing, and validation. We consider an array of four Imaging Cherenkov Telescopes (IACTs) and simulate observation of a fast-rotating star. The image predicted by the trained C-GAN shows promising results in reconstructing the star’s shape and size. The reconstructed brightness distributions are then assessed using moments.

>>>>

The foundational basis of II stems from the pioneering experiments and theoretical investigations, widely referred to as the \textquotedblleft HBT effect \textquotedblright, carried out initially by Hanbury Brown and Twiss (HBT) \cite{HBT56Lab,brown1957interferometry, brown1958interferometry} and, later, by Glauber \citep{glauber1963quantum}. The correlation between photons (called HBT effect) measured by a pair of photon detectors in two coherent beams of light was reported in 1956 \cite{HBT56Lab}. \cite{Rai2025} present a modern-day variant of this experiment, carried out with pseudo-thermal light. The HBT Effect and the related theoretical investigations laid the foundation for the modern field of Quantum Optics.

@@

As it stands today, studies grappling various issues of image reconstruction are being reported \citep{Haubois2009, Norris2021AZCyg, Liu2024SuperresolutionII, Liu2025}. MI-based image reconstruction has made substantial progress in this area with efforts at generating constructed images of stars like Betelgeuse \citep{Haubois2009} and AZ Cyg \citep{Norris2021AZCyg}. On the other hand, II-based methods are in a nascent stage. The recent publications \cite{Liu2024SuperresolutionII, Liu2025} have demonstrated, through outdoor experiments, imaging millimeter-scale targets at 1.36 km with a resolution ~14 times better than a single telescope’s diffraction limit. A \textquotedblleft flexible computational algorithm \textquotedblright reconstructs images from intensity correlations, overcoming atmospheric turbulence and optical imperfections.

We report here, the results of our attempt, the first of its kind, to reconstruct the II-simulated gravity-darkened images of fast rotating stars using a C-GAN neural network architecture. Image reconstruction in gravity darkened fast rotating stars has long been examined using various methods in MI \citep{vanBelle2001, DomicianodeSouza2003, DomicianodeSouza2005, mcalister2005first, Monnier2007, Pedretti2009, Zhao2009, Martinez2021}. Recently photosphere oblateness of $\gamma$-Cassiopeia \citep{Archer-arXiv-2025} has been measured at the VERITAS observatory using II. These results, and especially, that by \cite{Archer-arXiv-2025} put our work in context. Thus our simulation is a natural next step of the work of Archer et al and others. We emulate the C-GAN model \cite{isola2017image} to reconstruct images of fast-rotating stars using their simulated Intensity Interferograms and simulated sky-intensity distributions as input data for training, testing, and validation. We consider an array of four Imaging Cherenkov Telescopes (IACTs) and simulate observation of a fast-rotating star. The image predicted by the trained C-GAN shows promising results in reconstructing the star’s shape and size. The reconstructed brightness distributions are then assessed using moments.
